### 2020-05-28 writeup/transfer.tex
fix link

\textit{A simplifying setting.}
	Consider a setting where $\Sigma_1 = \Sigma_2 = \id$, in other words there is no covariate shift between the two tasks. For the noises, we assume that $\e_i$ has i.i.d. entries with mean zero and variances $\sigma_i^2$, $i=1,2$.
	For the task models, suppose that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ also has i.i.d. entries with mean $0$ and variance $d^2$.
	%The noise level for both tasks is equal to $\sigma > 0$.
	We have $n_i = \rho_i \cdot p$ data points from each task, for $i = 1, 2$.
	For simplicity, we assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in the supplementary material.

@@

\begin{proposition}\label{prop_dist_transition}
	In the simplifying setting with $\sigma_1 = \sigma_2 = \sigma$, assume that $\rho_1 > 40$ is a fixed constant.
	Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between the model distance and the noise level:
	\begin{itemize}
		\item If ${d^2} < \frac{1}{2}\cdot \frac {\sigma^2} {p} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If ${d^2} \ge 2\cdot\frac {\sigma^2} {p} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}

@@

\subsection{Improving Labeled Data Efficiency}\label{sec_data_size}

@@

\begin{proposition}\label{prop_data_size}
	In the setting of Example \ref{ex_basic}, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants.
	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
\item If $d^2 \le \frac{\sigma^2}{2p(\rho_2-1)}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.

@@

\begin{example}\label{ex_cov_family}
	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, we assume an upper bound on the scale of $M$.
	Consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	For the task models, we assume that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ has i.i.d. entries with mean $0$ and variance $d^2$.
	The following proposition shows that when $n_1$ is large enough compared to $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized approximately within the family of $\cS_{\mu}$ when $M = \mu\id$.
\end{example}

\begin{proposition}\label{prop_covariate}
	In the setting of Example \ref{ex_basic}, assume that $\rho_1 > 3$ and $\rho_2>3$,
and $\|\Sigma_1\|\le C_1$ for some constant $C_1>0$. %As a parameter of $M \in \cS_{\mu}$,
The test error $\te(\hat{\beta}_t^{\MTL})$, as a function of $M\in \cS_{\mu}$, is minimized at the identity matrix $\mu\id$ approximately by
$1+ \bigo{\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}}$.
%\be\label{formular_covariate0} (\te(\hat{\beta}_t^{\MTL}))(\mu \id) \le \left[1+ C\left(\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}\right)\right]\cdot\min_{M\in \cal S_{\mu}}(\te(\hat{\beta}_t^{\MTL}))(M) ,\ee
%where the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $C_1$, but otherwise does not depend on $c_1$ and $c_2$.
	%is minimized when $M = \mu\id$.
\end{proposition}
Proposition \ref{prop_covariate} shows that when $n_1\gg n_2$, $te(\hat{\beta}_t^{\MTL})$ is minimized when $\Sigma_1$ and $\Sigma_2$ are proportional to each other.
In other words, there is no covariate shift between the source task data and target task data.
This provides evidence that covariate shift is unfavorable when there are many source task datapoints,
To complement the result, we show an example when the statement is not true if $n_1 \le n_2$.

@@

\begin{proposition}\label{prop_var_transition}
	In the setting of Example \ref{ex_basic} with $d$ being fixed but $\sigma_1$ varies, assume that $\rho_1 > 50$ is a fixed constant and $d^2 < \frac {\sigma_2^2} {2p} \cdot \Phi(\rho_1, \rho_2)$.
	We derive the following transition as a parameter of $\sigma_1^2$:
	\begin{itemize}
		\item If $\sigma_1^2 \le -\frac32 \rho_1 \cdot p d^2+\left(1+ \frac34\rho_1 \Phi(\rho_1, \rho_2)\right)\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > - \frac12\rho_1\cdot p d^2   +\left(1+ \frac32\rho_1\Phi(\rho_1, \rho_2)\right) \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}
As a corollary, if $\sigma_1^2 \le \sigma_2^2$, then we always get positive transfer.
The proof of Proposition \ref{prop_var_transition} is similar to Proposition \ref{prop_dist_transition}.
The details can be found in Appendix \ref{app_proof_31}.

@@

\begin{proposition}\label{prop_data_efficiency}
	In the setting of Example \ref{ex_basic}, assume that $\rho_1 = \rho_2 = \rho \ge 200$ and $d^2 < {8\sigma^2} /{(3p \rho)}$.
	Then the data efficiency ratio is at most $\frac{1}{2\rho} + \frac{\sigma^2}{2\sigma^2 - 3p d^2 \rho / 4}$.
\end{proposition}
Note that we have stated the result assuming that $\rho_1 = \rho_2$.
Similar results can also be obtained when they are different.
We omit the details.
The proof of Proposition \ref{prop_data_efficiency} can be found in Appendix \ref{app_proof_data}.

>>>>

\textit{A simplified model.}
	Consider a setting where $\Sigma_1 = \Sigma_2 = \id$, in other words there is no covariate shift between the two tasks. For the noises, we assume that $\e_i$ has i.i.d. entries with mean zero and variances $\sigma_i^2$, $i=1,2$.
	For the task models, suppose that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ also has i.i.d. entries with mean $0$ and variance $d^2$.
	%The noise level for both tasks is equal to $\sigma > 0$.
	We have $n_i = \rho_i \cdot p$ data points from each task, for $i = 1, 2$.
	For simplicity, we assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in the supplementary material.

@@

\begin{proposition}\label{prop_dist_transition}
	In the simplified model, assume that $\rho_1 > 40$ is a fixed constant.
	Let $\sigma_1 = \sigma_2 = \sigma$.
	Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between the model distance and the noise level:
	\begin{itemize}
		\item If ${d^2} < \frac{1}{2}\cdot \frac {\sigma^2} {p} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If ${d^2} \ge 2\cdot\frac {\sigma^2} {p} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}

@@

@@

\begin{proposition}\label{prop_data_size}
	In the simplified model, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants.
	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
\item If $d^2 \le \frac{\sigma^2}{2p(\rho_2-1)}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.

@@

	To compare different choices of $M$ on the performance of $\hat{\beta}_t^{\MTL}$, we assume an upper bound on the scale of $M$.
	Consider the following family of matrices
	\begin{align*}
		\cS_{\mu}\define\bigset{M \mid \det\bigbrace{ M^\top M} \le \mu^p, \lambda(M) \in [\mu_{\min}, \mu_{\max}]},
	\end{align*}
	where $\mu, \mu_{\min}, \mu_{\max}$ are fixed values that do not grow with $p$.
	For the task models, we assume that $\beta_2$ has i.i.d. entries with mean zero and variance $\kappa^2$ and $\beta_1 - \beta_2$ has i.i.d. entries with mean $0$ and variance $d^2$.
	The following proposition shows that when $n_1$ is large enough compared to $n_2$, $\te(\hat{\beta}^{\MTL})$ is minimized approximately within the family of $\cS_{\mu}$ when $M = \mu\id$.

\begin{proposition}\label{prop_covariate}
	In the simplified model, assume that $\rho_1 > 3$ and $\rho_2>3$,
and $\|\Sigma_1\|\le C_1$ for some constant $C_1>0$. %As a parameter of $M \in \cS_{\mu}$,
The test error $\te(\hat{\beta}_t^{\MTL})$, as a function of $M\in \cS_{\mu}$, is minimized at the identity matrix $\mu\id$ approximately by
$1+ \bigo{\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}}$.
%\be\label{formular_covariate0} (\te(\hat{\beta}_t^{\MTL}))(\mu \id) \le \left[1+ C\left(\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}\right)\right]\cdot\min_{M\in \cal S_{\mu}}(\te(\hat{\beta}_t^{\MTL}))(M) ,\ee
%where the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $C_1$, but otherwise does not depend on $c_1$ and $c_2$.
	%is minimized when $M = \mu\id$.
\end{proposition}
Proposition \ref{prop_covariate} shows that when $n_1\gg n_2$, $te(\hat{\beta}_t^{\MTL})$ is minimized when $\Sigma_1$ and $\Sigma_2$ are proportional to each other.
In other words, there is no covariate shift between the source task data and target task data.
This provides evidence that covariate shift is unfavorable when there are many source task datapoints,
To complement the result, we show an example when the statement is not true if $n_1 \le n_2$.

@@

\begin{proposition}\label{prop_var_transition}
	In the simplified model with $d$ being fixed but $\sigma_1$ varies, assume that $\rho_1 > 50$ is a fixed constant and $d^2 < \frac {\sigma_2^2} {2p} \cdot \Phi(\rho_1, \rho_2)$.
	We derive the following transition as a parameter of $\sigma_1^2$:
	\begin{itemize}
		\item If $\sigma_1^2 \le -\frac32 \rho_1 \cdot p d^2+\left(1+ \frac34\rho_1 \Phi(\rho_1, \rho_2)\right)\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > - \frac12\rho_1\cdot p d^2   +\left(1+ \frac32\rho_1\Phi(\rho_1, \rho_2)\right) \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}
As a corollary, if $\sigma_1^2 \le \sigma_2^2$, then we always get positive transfer.
The proof of Proposition \ref{prop_var_transition} is similar to Proposition \ref{prop_dist_transition}.
The details can be found in Appendix \ref{app_proof_31}.

@@

\begin{proposition}\label{prop_data_efficiency}
	In the simplified model, assume that $\rho_1 = \rho_2 = \rho \ge 200$ and $d^2 < {8\sigma^2} /{(3p \rho)}$.
	Then the data efficiency ratio is at most $\frac{1}{2\rho} + \frac{\sigma^2}{2\sigma^2 - 3p d^2 \rho / 4}$.
\end{proposition}
Note that we have stated the result assuming that $\rho_1 = \rho_2$.
Similar results can also be obtained when they are different.
We omit the details.
The proof of Proposition \ref{prop_data_efficiency} can be found in Appendix \ref{app_proof_data}.
